Moral Foundations of U.S. Political News Organizations

### 2019-01-26 pilot_manuscript.Rmd
new Kohlberg and Gilligan stuff
based on their words

As a discipline, modern moral psychology started in the late 1960s with Lawrence Kohlberg [@Haidt2007]. Kohlberg's research popularized his theory of the development of moral reasoning. This theory establishes the steps of moral reasoning through which humans proceed as they mature. Kohlberg borrowed from Jean Piaget's stages of cognitive development in which children progress from the sensorimotor through to the formal operations stage [@Haidt2007]. Similarly, Kohlberg found people typically start with a "pre-conventional" understanding of morality during infancy and progress through several steps until they reach a "post-conventional" ethics. People who have reached the post-conventional stage are said to be able to weigh competing abstractions and reason their way to a conclusion that promotes justice. From Kohlberg's perspective, issues of justice and fairness comprise the foundation of morality [@Haidt2007]. This view persisted until it encountered criticism in the early 1980s.

Kohlberg's conception of morality faced major scrutiny from psychologist Carol Gilligan. In 1982, Gilligan criticized Kohlberg's theory on the grounds that it focused solely on the moral concerns of men, and that it ignored those of women [@Haidt2007]. She offered a historic argument contending women have traditionally filled roles related to caring and nurturing. Gilligan thus asserted morality was built upon an alternative moral foundation: caring [@Haidt2007]. This debate between competing conceptions of morality did not resolve until Gilligan and Kohlberg conceded the existence of two moral foundations: justice and caring [@Haidt2007]. While this new direction in moral psychology appeared to represent a more inclusive outlook on the construct, these novel ideas would soon be challenged on the grounds of its apparent western-centric outlook.

>>>>

As a discipline, modern moral psychology started in the late 1960s with Lawrence Kohlberg [@Haidt2007]. Kohlberg's research popularized his theory of the development of moral reasoning. This theory establishes the steps of moral reasoning through which humans proceed as their cognitive structures assume higher levels of sophistication and nuance [@Kohlberg1977]. Kohlberg borrowed from Jean Piaget's stages of cognitive development in which children progress from the sensorimotor through to the formal operations stage. Similarly, Kohlberg found people typically start with a "pre-conventional" understanding of morality during infancy in which children understand "right" and "wrong" purely in terms of how they interact with resultant experiences of rewards and punishment. Typically, people progress through several steps until they reach a "post-conventional" ethics. People who have reached the post-conventional stage are said to be able to weigh competing abstractions and reason their way to a conclusion that promotes justice based upon their "self-chosen ethical principles" [@Kohlberg1977]. From Kohlberg's perspective, issues of justice and fairness comprise the foundation of morality [@Haidt2007]. This view persisted until it encountered criticism in the early 1980s.

Kohlberg's conception of morality faced major scrutiny from psychologist Carol Gilligan. In 1982, Gilligan criticized Kohlberg's theory on the grounds that it focused solely on the moral concerns of men, and that it ignored those of women [@Haidt2007]. Gilligan drew attention to purported differences in the ways men and women are taught to relate to self and others. She offered a historic argument contending women have traditionally filled roles related to caring and nurturing. She pushes back against Kohlberg's assumption that moral development replaces "rule of brute force," as enforced by men, with the justice-based "rule of law." According to Gilligan, this assumption implies women are less morally developed, owing to their absence both in masculine displays of violence as well as in enforcement of the law [@Gilligan1982]. Gilligan argues for the existence of a distinct, but equal development process that women and girls must undergo in order to develop their moral selves. Stark differences in the ways women are traditionally taught to interact with their social world cause them to develop ethical systems based upon their non-aggressive relationships with others. Gilligan thus asserted morality was built upon an alternative moral foundation: caring [@Gilligan1982]. This debate between competing conceptions of morality did not resolve until Gilligan and Kohlberg conceded the existence of two moral foundations: justice and caring [@Haidt2007]. While this new direction in moral psychology appeared to represent a more inclusive outlook on the construct, these novel ideas would soon be challenged on the grounds of its apparent western-centric outlook.
